---
output:
  pdf_document:
    latex_engine: pdflatex
    citation_package: natbib
    template: mk_Rpackage_template.tex
    toc: true
    toc_depth: 2
    includes:
        in_header: vignette_head.tex
    keep_tex: true
title: "Honest inference in Regression Discontinuity Designs"
author: "Michal Kolesár"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 10pt
bibliography: np-testing-library.bib
vignette: >
  %\VignetteIndexEntry{Honest inference in Sharp Regression Discontinuity}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE, cache=FALSE}
library("knitr")
knitr::opts_knit$set(self.contained = FALSE)
knitr::opts_chunk$set(tidy = TRUE, collapse=TRUE, comment = "#>",
                      tidy.opts=list(blank=FALSE, width.cutoff=55))
```

# Introduction

The package `RDHonest` implements confidence intervals for the regression
discontinuity parameter considered in @ArKo16optimal, @ArKo16honest, and
@KoRo16. In this vignette, we demonstrate the implementation of these confidence
intervals using datasets from @lee08, @oreopoulos06, and @battistin09, which are
included in the package as a data frame `lee08`, `cghs`, and `rcp`. The datasets
from @lalive08 and @LuMi07 that are used in @ArKo16honest, and @KoRo16 are also
included in the package as data frames `rebp` and `headst`.

# Sharp RD

## Model

In the sharp regression discontinuity model, we observe units $i=1,\dotsc,n$,
with the outcome $y_i$ for the $i$th unit given by $$ y_i = f(x_i) + u_i, $$
where $f(x_i)$ is the expectation of $y_i$ conditional on the running variable
$x_i$ and $u_i$ is the regression error. A unit is treated if and only if the
running variable $x_{i}$ lies weakly above a known cutoff $x_{i}\geq c_{0}$. The
parameter of interest is given by the jump of $f$ at the cutoff, $$
\beta=\lim_{x\downarrow c_{0}}f(x)-\lim_{x\uparrow c_{0}}f(x).$$ Let
$\sigma^2(x_i)$ denote the conditional variance of $u_i$.

In the @lee08 dataset, the running variable corresponds to the margin of victory of
a Democratic candidate in a US House election, and the treatment corresponds to
winning the election. Therefore, the cutoff is zero. The outcome of interest is
the Democratic vote share in the following election.

The Oreopoulos dataset consists of a subsample of British workers, and it
exploits a change in minimum school leaving age in the UK from 14 to 15, which
occurred in 1947. The running variable is the year in which the individual turned
14, with the cutoff equal to 1947 so that the "treatment" is being subject to a
higher minimum school-leaving age. The outcome is log earnings in 1998.

Some of the functions in the package require the data to be transformed into a
custom `RDData` format. This can be accomplished with the `RDData` function:

```{r}
library("RDHonest")
```

## Plots

The package provides a function `RDScatter` to plot the raw data. To remove
some noise, the function plots averages over `avg` number of observations. The
function takes an `RDData` object as an argument

```{r, fig.width=4.5, fig.height=3.5, fig.cap="Lee (2008) data"}
## plot 25-bin averages in for observations 50 at most points
## away from the cutoff. See Figure 1.
RDScatter(voteshare~margin, data=lee08, subset=abs(lee08$margin)<=50,
          avg=50, propdotsize=FALSE, xlab="Margin of victory",
          ylab="Vote share in next election")
```

The running variable in the Oreopoulos dataset is discrete. It is therefore
natural to plot the average outcome by each value of the running variable, which
is achieved using by setting `avg=Inf`. The option `dotsize="count"` makes the
size of the points proportional to the number of observations that the point
averages over.

```{r, fig.width=4.5, fig.height=3.5, fig.cap="Oreopoulos (2006) data"}
## see Figure 2
f2 <- RDScatter(I(log(earnings))~yearat14, data=cghs, cutoff=1947,
                avg=Inf, xlab="Year aged 14", ylab="Log earnings",
                propdotsize=TRUE)
## Adjust size of dots if they are too big
f2 + ggplot2::scale_size_area(max_size = 4)
```

## Inference based on local polynomial estimates

The function `RDHonest` constructs one- and two-sided confidence intervals (CIs)
around local linear and local quadratic estimators using either a user-supplied
bandwidth (which is allowed to differ on either side of the cutoff), or
bandwidth that is optimized for a given performance criterion. The sense of
honesty is that, if the regression errors are normally distributed with known
variance, the CIs are guaranteed to achieve correct coverage _in finite
samples_, and achieve correct coverage asymptotically uniformly over the
parameter space otherwise. Furthermore, because the CIs explicitly take into
account the possible bias of the estimators, the asymptotic approximation
doesn't rely on the bandwidth to shrink to zero at a particular rate.

To describe the form of the CIs, let $\hat{\beta}_{h_{+},h_{-}}$ denote a a
local polynomial estimator with bandwidth equal to $h_{+}$ above the cutoff and
equal to $h_{-}$ below the cutoff. Let $\beta_{h_{+},h_{-}}(f)$ denote its
expectation conditional on the covariates when the regression function equals
$f$. Then the bias of the estimator is given by $\beta_{h_{+},h_{-}}(f)-\beta$.
Let $$
B(\hat{\beta}_{h_{+},h_{-}})=\sup_{f\in\mathcal{F}}|\beta_{h_{+},h_{-}}(f)-\beta|
$$ denote the worst-case bias over the parameter space $\mathcal{F}$. Then the
lower limit of a one-sided CI is given by $$\hat{\beta}_{h_{+},h_{-}}-
B(\hat{\beta}_{h_{+},h_{-}})-z_{1-\alpha}\widehat{se}(\hat{\beta}_{h_{+},h_{-}}),
$$ where $z_{1-\alpha}$ is the $1-\alpha$ quantile of a standard normal
distribution, and $\widehat{se}(\hat{\beta}_{h_{+},h_{-}})$ is the standard
error (an estimate of the standard deviation of the estimator). Subtracting the
worst-case bias in addition to the usual critical value times standard error
ensures correct coverage at all points in the parameter space.

A two-sided CI is given by $$ \hat{\beta}_{h_{+},h_{-}} \pm
cv_{1-\alpha}(B(\hat{\beta}_{h_{+},h_{-}})/\widehat{se}(\hat{\beta}_{h_{+},h_{-}}))\times
\widehat{se}(\hat{\beta}_{h_{+},h_{-}}),$$ where the critical value function
$cv_{1-\alpha}(b)$ corresponds to the $1-\alpha$ quantile of the $|N(b,1)|$
distribution. To see why using this critical value ensures honesty, decompose
the $t$-statistic as $$
\frac{\hat{\beta}_{h_{+},h_{-}}-\beta}{\widehat{se}(\hat{\beta}_{h_{+},h_{-}})}
=
\frac{\hat{\beta}_{h_{+},h_{-}}-\beta_{h_{+},h_{-}}(f)}{\widehat{se}(\hat{\beta}_{h_{+},h_{-}})}
+\frac{\beta_{h_{+},h_{-}}(f)-\beta}{\widehat{se}(\hat{\beta}_{h_{+},h_{-}})} $$
By a central limit theorem, the first term on the right-hand side will by
distributed standard normal, irrespective of the bias. The second term is
bounded in absolute value by
$B(\hat{\beta}_{h_{+},h_{-}})/\widehat{se}(\hat{\beta}_{h_{+},h_{-}})$, so that,
in large samples, the $1-\alpha$ quantile of the absolute value of the
$t$-statistic will be bounded by
$cv_{1-\alpha}(B(\hat{\beta}_{h_{+},h_{-}})/\widehat{se}(\hat{\beta}_{h_{+},h_{-}}))$.
This approach gives tighter CIs than simply adding and subtracting
$B(\hat{\beta}_{h_+,h_-})$ from the point estimate, in addition to adding and subtracting $z_{1-\alpha}\widehat{se}(\hat{\beta}_{h_+,h_-})$

The function `CVb` gives these critical values:

```{r, }
CVb(0, alpha=0.05) ## Usual critical value
CVb(1/2, alpha=0.05)

## Tabulate critical values for different bias levels
CVb(0:5, alpha=0.1)
```

### Parameter space

To implement the honest CIs, one needs to specify the parameter space
$\mathcal{F}$. The function `RDHonest` computes honest CIs when the parameter
space $\mathcal{F}$ corresponds to a second-order Taylor or second-order Hölder
smoothness class, which capture two different types of smoothness restrictions.
The second-order Taylor class assumes that $f$ lies in the the class of
functions $$\mathcal{F}_{\text{Taylor}}(M)= \left \{f_{+}-f_{-}\colon
f_{+}\in\mathcal{F}_{T}(M; [c_{0}
%] %(
,\infty)),\;
f_{-}\in\mathcal{F}_{T}(M;(-\infty, c_{0})) \right\},$$

where $\mathcal{F}_{T}(M;\mathcal{X})$ consists of functions $f$ such that the
approximation error from second-order Taylor expansion of $f(x)$ about $c_{0}$
is bounded by $M|x|^{2}/2$, uniformly over $\mathcal{X}$: \begin{align*}
\mathcal{F}_{T}(M;\mathcal{X}) =\left\{f\colon \left|
f(x)-f(c_0)-f'(c_0)x\right| \leq M|x|^2/2\text{ all }x\in\mathcal{X}\right\}.
\end{align*} The class $\mathcal{F}_{T}(M;\mathcal{X})$ formalizes the idea that
the second derivative of $f$ at zero should be bounded by $M$. See Section 2 in
@ArKo16optimal (note the constant $C$ in that paper equals $C=M/2$ here). This
class is doesn't impose smoothness away from boundary, which may be undesirable
in some empirical applications. The Hölder class addresses this problem by
bounding the second derivative globally. In particular, it assumes that $f$ lies
in the class of functions $$\mathcal{F}_{\text{Hölder}}(M)= \left
\{f_{+}-f_{-}\colon f_{+}\in\mathcal{F}_{H}(M;[c_{0}%] %( ,\infty)),\;
f_{-}\in\mathcal{F}_{H}(M;(-\infty, c_{0})) \right\},$$

where $$ \mathcal{F}_{H}(M;\mathcal{X})=\{f\colon |f'(x)-f'(y)|\leq M|x-y|
\;\;x,y\in\mathcal{X}\}.$$

The smoothness class is specified using the option `sclass`. CIs around a local
linear estimator with bandwidth that equals to 10 on either side of the cutoff
when the parameter space is given by a Taylor and Hölder smoothness class,
respectively, with $M=0.1$:

```{r}
RDHonest(voteshare~margin, data=lee08, kern="uniform", M=0.1, h=10, sclass="T")
RDHonest(voteshare~margin, data=lee08, kern="uniform", M=0.1, h=10, sclass="H")
```

The confidence intervals use the nearest-neighbor method to estimate the
standard error by default (this can be changed using the option `se.method`, see
help file for `RDHonest`). The package reports two-sided as well one-sided CIs
(with lower as well as upper limit) by default.

Instead of specifying a bandwidth, one can just specify the smoothness class and
smoothness constant $M$, and the bandwidth will be chosen optimally for a given
optimality criterion:

```{r}
RDHonest(voteshare ~ margin, data=lee08, kern="triangular",
         M=0.1, opt.criterion="MSE", sclass="H")
## Choose bws optimal for length of CI
RDHonest(voteshare ~ margin, data=lee08, kern="triangular", M=0.1,
         opt.criterion="FLCI", sclass="H")
```

### Inference when running variable is discrete

The confidence intervals described above can also be used when the running
variable is discrete, with $G$ support points: their construction makes no
assumptions on the nature of the running variable (see Section 5.1 in @KoRo16
for more detailed discussion).

Note that units that lies exactly at the cutoff are considered treated, since
the definition of treatment is that the running variable
 $x_i\geq c_0$.

As an example, consider the @oreopoulos06 data, in which the running variable is age in years:
```{r}
## Replicate Table 2, column (10)
RDHonest(log(earnings) ~ yearat14, cutoff=1947,
         data=cghs, kern="uniform", M=0.04, opt.criterion="FLCI", sclass="H")
## Triangular kernel generally gives tigher CIs
RDHonest(log(earnings) ~ yearat14, cutoff=1947,
         data=cghs, kern="triangular", M=0.04, opt.criterion="FLCI", sclass="H")
```

In addition, the package provides function `RDHonestBME` that calculates honest
confidence intervals under the assumption that the specification bias at zero is
no worse at the cutoff than away from the cutoff as in Section 5.2 in @KoRo16.

```{r}
## Replicate Table 2, column (6), run local linear regression (order=1)
## with a uniform kernel (other kernels are not yet implemented)
RDHonestBME(log(earnings) ~ yearat14, cutoff=1947,
            data=cghs, h=3, order=1)
```


Let us describe the implementation of the variance estimator $\hat{V}(W)$ used
to construct the CI as described in in Section 5.2 in @KoRo16. Suppose the point
estimate is given by the first element of the regression of the outcome $y_i$ on
$m(x_i)$. For instance, local linear regression with uniform kernel and
bandwidth $h$ corresponds to $m(x)=I(|x|\leq h)\cdot(I(x>c_0),1,x, x\cdot
I(x>c_0))'$. Let $\theta=Q^{-1}E[m(x_i)y_i]$, where $Q=E[m(x_i)m(x_i)']$, denote
the estimand for this regression (treating the bandwidth as fixed), and let
$\delta(x)=f(x)-m(x)'\theta$ denote the specification error at $x$. The RD
estimate is given by first element of the least squares estimator
$\hat{\theta}=\hat{Q}^{-1}\sum_i m(x_i)y_i$, where $\hat{Q}=\sum_i
m(x_i)m(x_i)'$.

Let $w(x_i)$ denote a vector of indicator (dummy) variables for all support
points of $x_i$ within distance $h$ of the cutoff, so that $\mu(x_g)$, where
$x_g$ is the $g$th support point of $x_i$, is given by the $g$th element of the
regression estimand $S^{-1}E[w(x_i)y_i]$, where $S=E[w(x_i)w(x_i)']$. Let
$\hat{\mu}=\hat{S}^{-1}\sum_i w(x_i)y_i$, where $\hat{S}=\sum_i w(x_i)w(x_i)'$
denote the least squares estimator. Then an estimate of
$(\delta(x_1),\dotsc,\delta(x_G))'$ is given by $\hat{\delta}$, the vector with
elements $\hat{\mu}_g-x_g\hat{\theta}$.

By standard regression results, the asymptotic distribution of $\hat{\theta}$
and $\hat{\mu}$ is given by
\begin{equation*}
\sqrt{n}
\begin{pmatrix}
\hat{\theta}-\theta\\
\hat{\mu}-\mu
\end{pmatrix}\overset{d}{\to}
\mathcal{N}\left(
0,
\Omega
\right),
\end{equation*}
where
\begin{equation*}
\Omega=\begin{pmatrix}
Q^{-1}E[(\epsilon_i^2+\delta(x_i)^2)m(x_i)m(x_i)']Q^{-1}&
Q^{-1}E[\epsilon_i^2 m(x_i)w(x_i)']S^{-1}\\
S^{-1}E[\epsilon_i^2 w(x_i)m(x_i)']Q^{-1}&
S^{-1}E[\epsilon_i^2 w(x_i)w(x_i)']S^{-1}\\
\end{pmatrix}.
\end{equation*}

Let $\hat{u}_i$ denote the regression residual from the regression of $y_i$ on
$m(x_i)$, and let $\hat{\epsilon}_i$ denote the regression residuals from the
regression of $y_i$ on $w(x_i)$. Then a consistent estimator of the asymptotic
variance $\Omega$ is given by
\begin{equation*}
\hat{\Omega}=n\sum_i T_i T_i',
\qquad
T_i'=\begin{pmatrix}
\hat{u}_i m(x_i)'\hat{Q}^{-1}&
\hat{\epsilon}_i w(x_i)'\hat{S}^{-1}
\end{pmatrix}.
\end{equation*}

Note that the upper left block and lower right block correspond simply to the
Eicker-Huber-White estimators of the asymptotic variance of $\hat{\theta}$ and
$\hat{\mu}$. By the delta method, a consistent estimator of the asymptotic
variance of $(\hat\delta,\hat{\theta}_1)$ is given by
\begin{equation*}
    \hat{\Sigma}=
\begin{pmatrix}
-X & I\\
e_1'& 0\\
\end{pmatrix}\hat{\Omega}\begin{pmatrix}
-X & I\\
e_1'& 0\\
\end{pmatrix}',
\end{equation*}
where $X$ is a matrix with $g$th row equal to $x_g'$, and $e_1$ is the first
unit vector.

Recall that in the notation of @KoRo16, $W=(g^-,g^+,s^-,s^+)$, and $g^{+}$ and
$g^{-}$ are such that $x_{g^{-}}< c_0\leq x_{g^{+}}$, and
$s^{+},s^{-}\in\{-1,1\}$. An upper limit for a right-sided CI for
$\theta_1+b(W)$ is then given by

\begin{equation*}
\hat{\theta}_{1}+s^{+}\hat\delta(x_{g^+})+
s^{-}\hat\delta(x_{g^-})+z_{1-\alpha}\hat{V}(W),
\end{equation*}

where $\hat{V}(W)=a(W)'\hat{\Sigma}a(W)$, and $a(W)\in\mathbb{R}^{G_{h}+1}$
denotes a vector with the $g_{-}$th element equal to $s^{-}$,
$(G_{h}^{-}+g_{+})$th element equal to $s^{+}$, the last element equal to one,
and the remaining elements equal to zero. The rest of the construction then
follows the description in Section 5.2 in @KoRo16.

## Data-driven choice of smoothness constant

Without further restrictions, the smoothness constant $M$ cannot be data-driven:
to maintain honesty over the whole function class, a researcher must choose
$M$ a priori, rather than attempting to use a data-driven method. Therefore,
one should, whenever possible, use problem-specific knowledge to decide what
choice of $M$ is reasonable a priori.

For cases in which this is difficult, the internal function `RDHonest`
implements the method considered in Section 3.4.1 in @ArKo16honest based on a
global polynomial approximation:

```{r}
## Data-driven choice of M
RDHonest(voteshare ~ margin, data=lee08, kern="uniform", sclass="H",
         opt.criterion="MSE")
```
See @ArKo16honest for a discussion of the restrictions on the
parameter space under which this method yields honest inference.

## Optimal inference

For the second-order Taylor smoothness class, the function `RDHonest`, with
`kernel="optimal"`, computes finite-sample optimal estimators and confidence
intervals, as described in Section 2.2 in @ArKo16optimal. This typically yields
tighter CIs. Comparing the lengths of two-sided CIs with optimally chosen
bandwidths, using Silverman's rule of thumb to estimate the preliminary variance
estimate used to compute optimal bandwidths:

```{r}
r1 <- RDHonest(voteshare ~ margin, data=lee08, kern="optimal", M=0.1,
               opt.criterion="FLCI",
               se.method="nn")$coefficients
r2 <- RDHonest(voteshare ~ margin, data=lee08, kern="triangular", M=0.1,
               opt.criterion="FLCI", se.method="nn",
               sclass="T")$coefficients
r1$conf.high-r1$conf.low
r2$conf.high-r2$conf.low
```

## Specification testing

The package also implements lower-bound estimates for the smoothness constant
$M$ for the Taylor and Hölder smoothness class, as described in the supplements
to @KoRo16 and @ArKo16optimal

```{r}
r1 <- RDHonest(voteshare ~ margin, data=lee08, M=0.1, se.method="nn")
### Only use three point-average for averages of a 100 points closest to cutoff,
### and report results separately for points above and below cutoff
RDSmoothnessBound(r1, s=100, separate=TRUE, multiple=FALSE, sclass="T")

## Pool estimates based on observations below and above cutoff, and
## use three-point averages over the entire support of the running variable
RDSmoothnessBound(r1, s=100, separate=FALSE, multiple=TRUE, sclass="H")
```

## Weighted regression

In some cases, data is only observed as cell averages. For instance, suppose
that instead of observing the original `cghs` data, we only observe averages for
cells as follows:

```{r}
d <- cghs
## Make 20 groups based on observation number
d$mod <- seq_along(d$yearat14) %% 20
## Make cells defined as intersection of group and year
d$cell <- d$mod/100+d$yearat14
## Data with cell averages
dd <- data.frame()
for (j in unique(d$cell)){
    dd <- rbind(dd, data.frame(y=mean(log(d$earnings)[d$cell==j]),
                               x=mean(d$yearat14[d$cell==j]),
                               weights=length(d$yearat14[d$cell==j])))
}
```

The column `weights` gives the number of observations that each cell averages
over. In this case, if we weight the observations using `weights`, we can
recover the original estimates (and the same worst-case bias):

```{r}
RDHonest(log(earnings)~yearat14, cutoff=1947, h=5, data=cghs, M=0.1,
         se.method="nn")
RDHonest(y~x, cutoff=1947, weights=weights, h=5, data=dd, M=0.1,
         se.method="nn")
```

Note the variance estimates don't quite match, since the variance estimator is
different, but the worst-case bias and the point estimate are identical.

# Initial bandwidth estimation

We use @FaGi96 rule of thumb and @ImKa12 bandwidth selectors for preliminary
variance estimation.


# Fuzzy RD

## Model

In a fuzzy RD design, units are assigned to treatment if their running variable
$x_{i}$ weakly exceeds a cutoff $x_i\geq c_{0}$. However, the actual treatment
$d_{i}$ does not perfectly comply with the treatment assignment. Instead, the cutoff
induces a jump in the treatment probability. The resulting reduced-form and
first-stage regressions are given by
\begin{align*}
  y_{i}&=f_{1}(x_{i})+u_{i1}, &  d_{i}&=f_{2}(d_{i})+u_{i2},
\end{align*}
See Section 3.3 in @ArKo16honest for a more detailed description.

In the @battistin09 dataset, the treatment variable is an indicator for
retirement, and the running variable is number of years since being eligible to
retire. The cutoff is $0$. Individuals exactly at the cutoff are dropped from
the dataset. If there were individuals exactly at the cutoff, they are assumed
to be assigned to the treatment group.

Similarly to the `RDData` function, the `FRDData` function transforms the data
into an appropriate format:

```{r}
## Assumes first column in the data frame corresponds to outcome,
##  second to the treatment variable, and third to the running variable
## Outcome here is log of non-durables consumption
## dr <- FRDData(cbind(logf=log(rcp[, 6]), rcp[, c(3, 2)]), cutoff=0)
```

## Inference based on local polynomial estimates

The function `RDHonest` constructs one- and two-sided confidence intervals
(CIs) around local linear and local quadratic estimators using either a
user-supplied bandwidth (which is allowed to differ on either side of the
cutoff), or bandwidth that is optimized for a given performance criterion.

### Parameter space and initial estimate

To implement the honest CIs, one needs to specify the parameter space
$\mathcal{F}$ for $f_1$ and $f_2$. The function `RDHonest` computes honest CIs
when $f_1$ and $f_2$ both lie in a second-order Taylor or second-order Hölder
smoothness class, $\mathcal{F}_{T}(M_1, M_2)$ and
$\mathcal{F}_{\text{Hölder}}(M_1, M_2)$, where the smoothness constants $M_1$
and $M_2$ for the reduced form and the first stage are allowed to differ. Also,
since the worst-case bias calculation requires an estimate of the treatment
effect, for optimal bandwidth calculations, the user needs to supply an initial
estimator of the treatment effect


```{r}
## Initial estimate of treatment effect for optimal bandwidth calculations
r <- RDHonest(log(cn) ~ retired | elig_year, data=rcp, kern="triangular",
              M=c(0.001, 0.002), opt.criterion="MSE", sclass="H", T0=0)
## Use it to compute optimal bandwidth
RDHonest(log(cn) ~ retired | elig_year, data=rcp, kern="triangular",
         M=c(0.001, 0.002), opt.criterion="MSE", sclass="H",
         T0=r$coefficients$estimate)
```

## Data-driven choice of smoothness constant

Like in the sharp RD case, Without further restrictions, the smoothness
constants $M_1$ and $M_2$ cannot be data-driven: to maintain honesty over the
whole function class, a researcher must choose them a priori, rather than
attempting to use a data-driven method. Therefore, one should, whenever
possible, use problem-specific knowledge to decide what choices of $M_1$ and
$M_2$ are reasonable a priori.

For cases in which this is difficult, the function `RDHonest` implements the
method considered in Section 3.4.1 in @ArKo16honest based on a global polynomial
approximation:

```{r}
## Data-driven choice of M
RDHonest(log(cn) ~ retired | elig_year, data=rcp, kern="triangular",
         opt.criterion="MSE", sclass="H", T0=r$coefficients$estimate)
```
See @ArKo16honest for a discussion of the restrictions on the
parameter space under which this method yields honest inference.

# Inference at a point

The package can also perform inference at
a point, and optimal bandwidth selection for inference at a point. Suppose, for example, one
was interested in the vote share for candidates with margin of victory equal to
20 points:
```{r}
## Transform data, specify we're interested in inference at x0=20,
## and drop observations below cutoff
RDHonest(voteshare ~ margin, data=lee08, subset=margin>0,
         cutoff=20, kern="uniform",
         opt.criterion="MSE", sclass="H", point.inference=TRUE)
```

# Lindeberg weights

In fuzzy RD, by Theorem B.2 in the appendix to @ArKo16honest, the estimator is asymptotically equivalent to
\begin{equation*}
\sum_i w_i (Y_i-D_i\beta)/first stage,
\end{equation*}
so the Lindeberg weights are the same as in the sharp case.


# References
