---
title: "Equivalent kernels and their moments"
author: "Michal Kolesár"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: false
vignette: >
  %\VignetteIndexEntry{Equivalent kernels and their moments}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

\newcommand{\abs}[1]{\lvert#1\rvert}

This package also computes equivalent kernels for local linear regression and
their moments. For orders $r=1,2$ (local linear and local quadratic), selected
moments and equivalent kernels are computed analytically for the the uniform,
triangular and Epanechnikov kernels. Higher-order moments, and moments for other
kernels with bounded support normalized to $[-1,1]$ are computed numerically.

The uniform, triangular and Epanechnikov kernels are defined as $$
k_{\text{U}}(u)=\frac{1}{2}I(|u|<1),\qquad k_{\text{T}}(u)=(1-|u|)I(|u|<1),
\qquad k_{\text{E}}(u)=\frac{3}{4}(1-u^2)I(|u|<1). $$

We compute the following moments
$$
\mu_{j}(k)=\int_{\mathcal{X}} u^j k(u) \, du,\qquad
\nu_{j}(k)=\int_{\mathcal{X}} u^j k^2(u) \, du,\qquad
\pi_{j}(k)=\int_{\mathcal{X}} |u^j k^2(u)| \, du,
$$
where $\mathcal{X}$ is $[0,1]$ for boundary regression, and $[-1,1]$ for
interior regression, and $k$ is a kernel of equivalent kernel.


# Equivalent kernels

Define $M_{r}(k)=\int_{\mathcal{X}}p_{r}(u)p_{r}(u)'k(u) \, du$, where
$p_{r}(u)=(1,u,\dotsc,u^{r})$, so that $(M_{r}(k))_{i,j}=\mu_{i+j-2}(k)$. Then for
local polynomial regression of order $r$ using a kernel $k$, the equivalent
kernel is given by
$$
k^*\colon u\mapsto e_{1}'M_{r}(k)^{-1}p(u)k(u), \qquad u\in\mathcal{X}
$$
where $e_{1}$ is a vector of zeros with $1$ in the first position. This assumes
that the parameter of interest is the value of the regression function $f$ at a
point, which we assume throughout. For $r=0$, the equivalent kernel is the same
as the original kernel, except it is normalized so that it integrates to one
over $\mathcal{X}$, $\int_{\mathcal{X}}k=1$.

For the three kernels and boundary regression of order $r=0,1,2$, the equivalent
kernels are the following functions with domain $[0,1]$

| Kernel | Order 0              | Order 1                       | Order 2                            |
|:-------|:---------------------|:------------------------------|:-----------------------------------|
| U      | $1$                  | $(4-6u)$                      | $3(3-12u+10u^2)$                   |
| T      | $2(1-u)$             | $6(1-2u)(1-u)$                | $12(1-5u+5u^2)(1-u)$               |
| E      | $\frac{3}{2}(1-u^2)$ | $\frac{12}{19}(8-15u)(1-u^2)$ | $\frac{5}{8}(17-80u+77u^2)(1-u^2)$ |

For interior regression, the equivalent kernels are the following functions with
domain $[-1,1]$ (orders 0 and 1 are the same)

| Kernel | Order 0 or 1         | Order 2                         |
|:-------|:---------------------|:--------------------------------|
| U      | $\frac{1}{2}$        | $\frac{1}{8}(9-15u^2)$          |
| T      | $(1-\abs{u})$        | $\frac{6}{7}(2-5u^2)(1-\abs{u})$|
| E      | $\frac{3}{4}(1-u^2)$ | $\frac{15}{32}(3-7u^2)(1-u^2)$  |

The equivalent kernels can be computed with the function `EqKern`, which returns the equivalent kernel function

```{r, echo=TRUE, results=FALSE}
library("RDHonest")
EqKern("uniform", boundary = TRUE, order = 2)(0.5)
# Equivalent call
EqKern(function(u) u <= 1, boundary = TRUE, order = 2 )(0.5)
```


# Kernel Moments

The package stores analytically-computed low-order moments for the uniform,
triangular, and Epanechnikov kernels for fast access in the dataframe `kernC`.
The moments for boundary kernels are as follows:

| Kernel       | Order | $\mu_0$ | $\mu_1$      | $\mu_2$        | $\mu_3$         | $\mu_4$           | $\nu_0$             | $\nu_1$           | $\nu_2$            | $\nu_3$            | $\nu_4$              |
|:-------------|:------|:--------|:-------------|---------------:|----------------:|------------------:|--------------------:|------------------:|-------------------:|-------------------:|---------------------:|
| U      |  0    |1        |$\frac{1}{2}$ |$\frac{1}{3}$   |$\frac{1}{4}$    |$\frac{1}{5}$      |$1$                  |$\frac{1}{2}$      |$\frac{1}{3}$       |$\frac{1}{4}$       |$\frac{1}{5}$         |
| T   |  0    |1        |$\frac{1}{3}$ |$\frac{1}{6}$   |$\frac{1}{10}$   |$\frac{1}{15}$     |$\frac{4}{3}$        |$\frac{1}{3}$      |$\frac{2}{15}$      |$\frac{1}{15}$      |$\frac{4}{105}$       |
| E |  0    |1        |$\frac{3}{8}$ |$\frac{1}{5}$   |$\frac{1}{8}$    |$\frac{3}{35}$     |$\frac{6}{5}$        |$\frac{3}{8}$      |$\frac{6}{35}$      |$\frac{3}{32}$      |$\frac{2}{35}$        |
| U      |  1    |1        |0             |$-\frac{1}{6}$  |$-\frac{1}{5}$   |$-\frac{1}{5}$     |$4$                  |$1$                |$\frac{8}{15}$      |$\frac{2}{5}$      |$\frac{12}{15}$       |
| T   |  1    |1        |0             |$-\frac{1}{10}$ |$-\frac{1}{10}$  |$-\frac{3}{35}$    |$\frac{24}{5}$       |$\frac{3}{5}$      |$\frac{6}{35}$      |$\frac{3}{35}$      |$\frac{2}{35}$        |
| E |  1    |1        |0             |$-\frac{11}{95}$|$-\frac{16}{133}$|$-\frac{141}{1330}$|$\frac{56832}{12635}$|$\frac{1770}{2527}$|$\frac{2976}{12635}$|$\frac{324}{2527}$  |$\frac{12352}{138985}$|
| U      |  2    |1        |0             |0               |$\frac{1}{20}$   |$\frac{3}{35}$     |$9$                  |$\frac{3}{2}$      |$\frac{27}{35}$     |$\frac{81}{140}$    |$\frac{17}{35}$       |
| T   |  2    |1        |0             |0               |$\frac{1}{35}$   |$\frac{3}{70}$     |$\frac{72}{7}$       |$\frac{6}{7}$      |$\frac{8}{35}$      |$\frac{4}{35}$      |$\frac{4}{35}$        |
| E |  2    |1        |0             |0               |$\frac{15}{448}$ |$\frac{13}{252}$   |$\frac{9895}{1008}$  |$\frac{4105}{4032}$|$\frac{325}{1008}$  |$\frac{2825}{16128}$|$\frac{16795}{144144}$|


For interior regression (recall order 0 and 1 give the same kernel)

| Kernel       | Order | $\mu_0$ | $\mu_2$      | $\mu_4$         | $\nu_0$         | $\nu_2$           | $\nu_4$            |
|:-------------|:------|:--------|:-------------|----------------:|----------------:|------------------:|-------------------:|
| U      |  0    |1        |$\frac{1}{3}$ |$\frac{1}{5}$    |$\frac{1}{2}$    |$\frac{1}{6}$      |$\frac{1}{10}$      |
| T   |  0    |1        |$\frac{1}{6}$ |$\frac{1}{15}$   |$\frac{2}{3}$    |$\frac{1}{15}$     |$\frac{2}{105}$     |
| E |  0    |1        |$\frac{1}{5}$ |$\frac{3}{35}$   |$\frac{3}{5}$    |$\frac{3}{35}$     |$\frac{1}{35}$      |
| U      |  2    |1        |0             |$-\frac{3}{35}$  |$\frac{9}{8}$    |$-\frac{9}{56}$    |$-\frac{23}{280}$   |
| T   |  2    |1        |0             |$-\frac{19}{490}$|$\frac{456}{343}$|$-\frac{106}{1715}$|$-\frac{256}{18865}$|
| E |  2    |1        |0             |$-\frac{1}{21}$  |$\frac{5}{4}$    |$-\frac{25}{308}$  |$-\frac{85}{4004}$  |

We also store absolute moments. At the boundary:

| Kernel       | Order | $\pi_0$                  | $\pi_1$                | $\pi_2$                   | $\pi_3$                       | $\pi_4$           |
|:-------------|:------|-------------------------:|-----------------------:|--------------------------:|------------------------------:|------------------:|
| U      |  0    |1                         |$\frac{1}{2}$           |$\frac{1}{3}$              |$\frac{1}{4}$                  |$\frac{1}{5}$      |
| T   |  0    |1                         |$\frac{1}{3}$           |$\frac{1}{6}$              |$\frac{2}{10}$                 |$\frac{1}{15}$     |
| E |  0    |1                         |$\frac{3}{8}$           |$\frac{3}{5}$              |$\frac{1}{8}$                  |$\frac{3}{35}$     |
| U      |  1    |$\frac{5}{3}$             |$\frac{16}{27}$         |$\frac{59}{162}$           |$\frac{113}{405}$              |$\frac{857}{3545}$ |
| T   |  1    |$\frac{3}{2}$             |$\frac{3}{8}$           |$\frac{3}{16}$             |$\frac{1}{8}$                  |$\frac{3}{32}$     |
| E |  1    |$\frac{100483}{64125}$    |$\frac{702464}{1603125}$|$\frac{16520549}{72140625}$|$\frac{235792912}{1514953125}$ |$\frac{1792724653}{15149531250}$ |
| U      |  2    |$1+12\frac{\sqrt{6}}{5^2}$|$\frac{36\sqrt{6}}{5^3}$|$\frac{558\sqrt{6}}{5^5}$  |$\frac{1782\sqrt{6}}{5^6}+1/20$|$0.268931$         |
| T   |  2    |$1+\frac{2}{\sqrt{5}}$    |$\frac{24\sqrt{5}}{5^3}$|$\frac{12\sqrt{5}}{5^3}$   |$\frac{218\sqrt{5}}{4375}+1/35$|$0.102656$         |
| E |  2    |$2.0051585$               |$0.50792888$            |$0.26617935$               |$0.17770885$                   |$0.1321148$        |

In the interior:

| Kernel       | Order | $\pi_0$ | $\pi_1$      | $\pi_2$      | $\pi_3$      | $\pi_4$      |
|:-------------|:------|:--------|:-------------|:-------------|:-------------|:-------------|
| U      |  0    |1        |$\frac{1}{2}$ |$\frac{1}{3}$ |$\frac{1}{4}$ |$\frac{1}{5}$ |
| T   |  0    |1        |$\frac{1}{3}$ |$\frac{1}{6}$ |$\frac{1}{10}$|$\frac{1}{15}$|
| E |  0    |1        |$\frac{3}{8}$ |$\frac{1}{5}$ |$\frac{1}{8}$ |$\frac{3}{35}$|
| U      |  2    |1        |$0.4875$      |$0.27885480$  |$0.1975$      |$0.157419807$ |
| T   |  2    |1        |$0.31155912$  |$0.13986945$  |$0.08443054$  |$0.060140887$ |
| E |  2    |1        |$0.36033164$  |$0.17177502$  |$0.10671010$  |$0.077066194$ |

Finally, in the column `kernC$pMSE`, the package stores the optimal constant for
pointwise MSE optimal bandwidth (see Fan and Gijbels, page 67):
$$
\left(\frac{(p+1)!^2\nu_{0}}{2(p+1)\mu_{p+1}^2}\right)^{\frac{1}{2p+3}},
$$
where $p$ is the order of local polynomial.

For other kernels, the moments can be computed using the `KernMoment` function:
```{r}
## mu_1, should be 0
KernMoment(function(u) 4-6*u, moment = 1, boundary = TRUE, type = "raw")
## nu_1, should be 0
KernMoment(function(u) 4-6*u, moment = 1, boundary = TRUE, type = "raw2")
## pi_1, should be 16/27
KernMoment(function(u) 4-6*u, moment = 1, boundary = TRUE, type = "absolute")
```
